---
title: "2024-10-14_MakeLeafcutterContrasts"
output: html_document
date: '2024-10-14'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

## Intro

As one last attempt to make the beat beta scatter heatmap right, let's use leafcutter differential splicing to calculate dpsi, and compare to log2FC.

First, let's sysytematically make the contrast tables.

```{r}
library(tidyverse)
library(edgeR)

# Set theme
theme_set(
  theme_classic() +
  theme(text=element_text(size=16,  family="Helvetica")))

# I use layer a lot, to rotate long x-axis labels
Rotate_x_labels <- theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

samples <- read_tsv("../code/config/Cordoso_Moreira_SampleList.tsv")

Stages <- read_tsv("../data/Stages_AsIn_CordosoMoreira_Recoded.txt")
```

Now make dot plot of samples, to better understand how we can split the timecourse into two for simple contrasts

```{r, fig.height=8, fig.width=8}
samples %>%
  filter(`Used library?` %in% c("yes", "Yes")) %>%
  separate_rows(Tissue_ForDevelopementalAnalysis) %>%
  count(Ordinal_stage, ID_Species, Tissue_ForDevelopementalAnalysis) %>%
  inner_join(
    Stages %>%
      dplyr::select(Ordinal_stage, Marker, ID_Species=Species)
  ) %>%
  ggplot(aes(x=Ordinal_stage, y=Tissue_ForDevelopementalAnalysis)) +
  geom_point(aes(size=n)) +
  # geom_line() +
  geom_vline(data = . %>%
               filter(!is.na(Marker)) %>%
               distinct(Marker, Ordinal_stage, ID_Species),
             aes(color=Marker, xintercept=Ordinal_stage)
               ) +
  facet_wrap(~ID_Species, scales="free") +
  theme(legend.position='bottom')
```



Maybe we can try stages before onset meiosis, versus after birth. or first 5 versus last 5. or pre-birth vs post-birth.

1. ) Before meiosis vs after spermatogenesis
2. ) Before vs after birth
3. ) Before birth vs after spermatogenesis
4. ) first half vs second half within tissue species set
5. ) First 10 samples in timecourse vs last 10 samples tissue-species set

First let's make contrasts for #1 (Before meiosis vs after spermatogenesis)

```{r}
Stages %>%
  dplyr::select(Ordinal_stage, Marker, ID_Species=Species) %>%
  filter(!is.na(Marker)) %>%
  distinct() %>%
  arrange(ID_Species, Ordinal_stage)
  


Approach1.samples.grouped <- samples %>%
  filter(`Used library?` %in% c("yes", "Yes")) %>%
  separate_rows(Tissue_ForDevelopementalAnalysis) %>%
  left_join(
    Stages %>%
      dplyr::select(LeftMarkerStage=Ordinal_stage, Marker, ID_Species=Species) %>%
      filter(!is.na(Marker)) %>%
      distinct() %>%
      filter(Marker == "Onset meiosis oogenesis") %>%
      dplyr::select(-Marker)
  ) %>%
  left_join(
    Stages %>%
      dplyr::select(RightMarkerStage=Ordinal_stage, Marker, ID_Species=Species) %>%
      filter(!is.na(Marker)) %>%
      distinct() %>%
      filter(Marker == "Onset meiosis spermatogenesis") %>%
      dplyr::select(-Marker)
  ) %>%
  mutate(
    Group = case_when(
      Ordinal_stage <= LeftMarkerStage ~ "Early",
      Ordinal_stage >= RightMarkerStage ~ "Late",
      TRUE ~ "Middle"
    )
  ) %>%
  filter(Group %in% c("Early", "Late")) %>%
  mutate(Group = relevel(factor(Group), "Early")) %>%
  dplyr::select(Group, Tissue_ForDevelopementalAnalysis, ID_Species, ID)

Approach1.samples.grouped %>%
  count(ID_Species, Group, Tissue_ForDevelopementalAnalysis) %>%
  ggplot(aes(x=Tissue_ForDevelopementalAnalysis, y=n, fill=Group)) +
  geom_col(position='dodge') +
  facet_wrap(~ID_Species) +
  Rotate_x_labels
  
```

So we will obviously be missing some tissue:species pairs with this set up. but that's ok.

Let's make the other contrasts...

```{r}
Approach2.samples.grouped <- samples %>%
  filter(`Used library?` %in% c("yes", "Yes")) %>%
  separate_rows(Tissue_ForDevelopementalAnalysis) %>%
  left_join(
    Stages %>%
      dplyr::select(LeftMarkerStage=Ordinal_stage, Marker, ID_Species=Species) %>%
      filter(!is.na(Marker)) %>%
      distinct() %>%
      filter(Marker == "Birth") %>%
      dplyr::select(-Marker)
  ) %>%
  left_join(
    Stages %>%
      dplyr::select(RightMarkerStage=Ordinal_stage, Marker, ID_Species=Species) %>%
      filter(!is.na(Marker)) %>%
      distinct() %>%
      filter(Marker == "Birth") %>%
      dplyr::select(-Marker)
  ) %>%
  mutate(
    Group = case_when(
      Ordinal_stage < LeftMarkerStage ~ "Early",
      Ordinal_stage > RightMarkerStage ~ "Late",
      TRUE ~ "Middle"
    )
  ) %>%
  filter(Group %in% c("Early", "Late")) %>%
  mutate(Group = relevel(factor(Group), "Early")) %>%
  dplyr::select(Group, Tissue_ForDevelopementalAnalysis, ID_Species, ID)

Approach3.samples.grouped <- samples %>%
  filter(`Used library?` %in% c("yes", "Yes")) %>%
  separate_rows(Tissue_ForDevelopementalAnalysis) %>%
  left_join(
    Stages %>%
      dplyr::select(LeftMarkerStage=Ordinal_stage, Marker, ID_Species=Species) %>%
      filter(!is.na(Marker)) %>%
      distinct() %>%
      filter(Marker == "Birth") %>%
      dplyr::select(-Marker)
  ) %>%
  left_join(
    Stages %>%
      dplyr::select(RightMarkerStage=Ordinal_stage, Marker, ID_Species=Species) %>%
      filter(!is.na(Marker)) %>%
      distinct() %>%
      filter(Marker == "Onset meiosis spermatogenesis") %>%
      dplyr::select(-Marker)
  ) %>%
  mutate(
    Group = case_when(
      Ordinal_stage < LeftMarkerStage ~ "Early",
      Ordinal_stage >= RightMarkerStage ~ "Late",
      TRUE ~ "Middle"
    )
  ) %>%
  filter(Group %in% c("Early", "Late")) %>%
  mutate(Group = relevel(factor(Group), "Early")) %>%
  dplyr::select(Group, Tissue_ForDevelopementalAnalysis, ID_Species, ID)

Approach4.samples.grouped <- samples %>%
  filter(`Used library?` %in% c("yes", "Yes")) %>%
  separate_rows(Tissue_ForDevelopementalAnalysis) %>%
  group_by(ID_Species, Tissue_ForDevelopementalAnalysis) %>%
  mutate(Ordinal_Percent_Rank = percent_rank(Ordinal_stage)) %>%
  ungroup() %>%
  mutate(
    Group = case_when(
      Ordinal_Percent_Rank < 0.5 ~ "Early",
      Ordinal_Percent_Rank >= 0.5 ~ "Late",
      TRUE ~ "Middle"
    )) %>%
  filter(Group %in% c("Early", "Late")) %>%
  mutate(Group = relevel(factor(Group), "Early")) %>%
  dplyr::select(Group, Tissue_ForDevelopementalAnalysis, ID_Species, ID)

Approach5.samples.grouped <- samples %>%
  filter(`Used library?` %in% c("yes", "Yes")) %>%
  separate_rows(Tissue_ForDevelopementalAnalysis) %>%
  group_by(ID_Species, Tissue_ForDevelopementalAnalysis) %>%
  mutate(Ordinal_Rank = rank(Ordinal_stage, ties.method='random')) %>%
  mutate(Ordinal_RevRank = rank(desc(Ordinal_stage), ties.method='random')) %>%
  ungroup() %>%
  mutate(
    Group = case_when(
      Ordinal_Rank <= 10 & Ordinal_RevRank > 10 ~ "Early",
      Ordinal_RevRank <= 10 & Ordinal_Rank > 10 ~ "Late",
      TRUE ~ "Middle"
    )) %>%
  filter(Group %in% c("Early", "Late")) %>%
  mutate(Group = relevel(factor(Group), "Early")) %>%
  dplyr::select(Group, Tissue_ForDevelopementalAnalysis, ID_Species, ID) %>%
  arrange(ID_Species, Tissue_ForDevelopementalAnalysis, Group)

```



And now look at sample sizes in contrasts

```{r}
bind_rows(
  list(
    "BeforeM_vs_AfterS" = Approach1.samples.grouped,
    "BeforeB_vs_AfterB" = Approach2.samples.grouped,
    "BeforeB_vs_AferS" = Approach3.samples.grouped,
    "FirstHalf_vs_SecondHalf" = Approach4.samples.grouped,
    "First10_vs_Second10" = Approach5.samples.grouped
  ),
  .id="Approach"
) %>%
  count(Group, Tissue_ForDevelopementalAnalysis, Approach, ID_Species) %>%
  ggplot(aes(x=Tissue_ForDevelopementalAnalysis, y=n, fill=Group)) +
  geom_col(position='stack') +
  facet_grid(Approach~ID_Species) +
  Rotate_x_labels

Contrasts <- bind_rows(
  list(
    "BeforeM_vs_AfterS" = Approach1.samples.grouped,
    "BeforeB_vs_AfterB" = Approach2.samples.grouped,
    "BeforeB_vs_AferS" = Approach3.samples.grouped,
    "FirstHalf_vs_SecondHalf" = Approach4.samples.grouped,
    "First10_vs_Second10" = Approach5.samples.grouped
  ),
  .id="Approach"
) %>%
  add_count(Group, Tissue_ForDevelopementalAnalysis, Approach, ID_Species) %>%
  group_by(Group, Tissue_ForDevelopementalAnalysis, Approach, ID_Species) %>%
  filter(all(n>=2)) %>%
  ungroup() %>%
  group_by(Tissue_ForDevelopementalAnalysis, Approach, ID_Species) %>%
  filter(any(Group=="Late") & any(Group=="Early")) %>%
  ungroup() %>%
  mutate(GenomeName = recode(ID_Species, "Chicken"="Chicken_UCSC.galGal6_ensv101", "Rabbit"="Rabbit_UCSC.oryCun2_ensv101", "Opossum"="Opossum_UCSC.monDom5_ensvUnknown", "Human"="Human_UCSC.hg38_GencodeComprehensive46", "Mouse"="Mouse_UCSC.mm39_GencodeComprehensive46", "Rat"="Rat_UCSC.rn7_RefSeqv108", "Macaque"="Macaque_UCSC.rheMac10_ensv101")) %>%
  mutate(ContrastName = str_glue("{Approach}.{Tissue_ForDevelopementalAnalysis}.{ID_Species}"))


Contrasts %>%
  distinct(ContrastName, GenomeName) %>%
  dplyr::select(ContrastName, GenomeName) %>%
  write_tsv("../code/config/CordosoTimeSeriesContrasts.tsv")

dir.create("../code/MazinLeafcutterAnalysis/ContrastGroupFiles/")

Contrasts %>%
  dplyr::select(ContrastName, ID, ID_Species, Tissue_ForDevelopementalAnalysis, Group) %>%
  arrange(ID_Species, Tissue_ForDevelopementalAnalysis, Group) %>%
  group_by(ContrastName) %>%
  dplyr::select(ID, Group) %>%
  group_walk(~ write_delim(.x, paste0("../code/MazinLeafcutterAnalysis/ContrastGroupFiles/", .y$ContrastName, ".txt"), delim = ' ', col_names = F))
```

And now let's prototype a quick DE analysis from the same contrast file

```{r}
Example.contrast <- read_delim("../code/MazinLeafcutterAnalysis/ContrastGroupFiles/BeforeB_vs_AferS.Brain.Chicken.txt", col_names = c("ID", "Group"), delim = ' ')

Counts <- fread("../code/MazinLeafcutterAnalysis/Expression/Chicken_UCSC.galGal6_ensv101.counts.tsv.gz") %>%
  dplyr::select(Geneid, Example.contrast$ID) %>%
  column_to_rownames("Geneid")

y <- DGEList(counts=Counts, group=Example.contrast$Group)

y$samples


keep <- filterByExpr(y)
y <- y[keep, , keep.lib.sizes=FALSE]

dim(y)

results <- y %>%
  calcNormFactors() %>%
  estimateDisp() %>%
  exactTest() %>%
  topTags(n=Inf) %>%
  as.data.frame() %>%
  rownames_to_column("Geneid")

results %>%
  ggplot(aes(x=logFC, y=-log10(PValue), color=FDR<0.01)) +
  geom_point(alpha=0.1)

results %>%
  ggplot(aes(x=logCPM, y=logFC, color=FDR<0.01)) +
  geom_point(alpha=0.1)

results %>%
  mutate(IsSig = FDR<0.01) %>%
  count(IsSig)

```


Lastly, I also want to include a contrast for neonate vs mature adult (post spermatogenesis meiosis), and a contrast for each mature adult tissue pair within species.

```{r}
# How many post-birth samples are there for each tissue/species
samples %>%
  filter(`Used library?` %in% c("yes", "Yes")) %>%
  separate_rows(Tissue_ForDevelopementalAnalysis) %>%
  left_join(
    Stages %>%
      dplyr::select(LeftMarkerStage=Ordinal_stage, Marker, ID_Species=Species) %>%
      filter(!is.na(Marker)) %>%
      distinct() %>%
      filter(Marker == "Birth") %>%
      dplyr::select(-Marker)
  ) %>%
  filter(Ordinal_stage >= LeftMarkerStage) %>%
  arrange(ID_Species, Tissue_ForDevelopementalAnalysis, Ordinal_stage) %>%
  count(ID_Species, Tissue_ForDevelopementalAnalysis)


# How many neonate samples are there
samples %>%
  filter(`Used library?` %in% c("yes", "Yes")) %>%
  separate_rows(Tissue_ForDevelopementalAnalysis) %>%
  left_join(
    Stages %>%
      dplyr::select(LeftMarkerStage=Ordinal_stage, Marker, ID_Species=Species) %>%
      filter(!is.na(Marker)) %>%
      distinct() %>%
      filter(Marker == "Birth") %>%
      dplyr::select(-Marker)
  ) %>%
  filter(Ordinal_stage == LeftMarkerStage) %>%
  count(ID_Species, Tissue_ForDevelopementalAnalysis)

```

So for some species, we might need to broaden the definition of neonate to include all tissues

```{r}
samples %>%
  filter(`Used library?` %in% c("yes", "Yes")) %>%
  separate_rows(Tissue_ForDevelopementalAnalysis) %>%
  left_join(
    Stages %>%
      dplyr::select(LeftMarkerStage=Ordinal_stage, Marker, ID_Species=Species) %>%
      filter(!is.na(Marker)) %>%
      distinct() %>%
      filter(Marker == "Birth") %>%
      dplyr::select(-Marker)
  ) %>%
  filter(Ordinal_stage >= LeftMarkerStage) %>%
  arrange(ID_Species, Tissue_ForDevelopementalAnalysis, Ordinal_stage) %>%
  add_count(Tissue_ForDevelopementalAnalysis, ID_Species, Ordinal_stage) %>%
  distinct(Tissue_ForDevelopementalAnalysis, ID_Species, Ordinal_stage, .keep_all=T) %>%
  mutate(label = str_glue("{Label_As_in_PreviousAnalysis};{n}")) %>%
  ggplot(aes(x=Ordinal_stage, y=Tissue_ForDevelopementalAnalysis)) +
    geom_point(aes(size=n)) +
    # geom_text(aes(label=label)) +
    facet_wrap(~ID_Species, scales="free_x")
  
```
Ok I think the best will be manual labelling neonate groups for each tissue in excel... Let's write out a tsv file to open in excel, then I'll manually make neonate and adult tissue groups

```{r}
samples %>%
  filter(`Used library?` %in% c("yes", "Yes")) %>%
  separate_rows(Tissue_ForDevelopementalAnalysis) %>%
  left_join(
    Stages %>%
      filter(!is.na(Marker)) %>%
      dplyr::select(Ordinal_stage, Marker, ID_Species=Species) %>%
      distinct()
  ) %>%
  arrange(ID_Species, Tissue_ForDevelopementalAnalysis, Ordinal_stage) %>%
  dplyr::select(ID, ID_Species, Tissue_ForDevelopementalAnalysis, Ordinal_stage, Label_As_in_PreviousAnalysis, Marker) %>%
  write_tsv("../code/scratch/Samples_ForManualNeonateAndAdultClassification.tsv")
```

After manual annotation...

```{r}
ManuallyAnnotated_Neonate_v_Adults <- read_tsv("../code/scratch/Samples_ForManualNeonateAndAdultClassification_ManuallyClassified.txt")

ManuallyAnnotated_Neonate_v_Adults %>%
  filter(!is.na(DifferentialSplicingAndExpressionGroup)) %>%
  add_count(Ordinal_stage, Tissue_ForDevelopementalAnalysis, ID_Species) %>%
  distinct(Ordinal_stage, Tissue_ForDevelopementalAnalysis, ID_Species, .keep_all = T) %>%
  ggplot(aes(x=Ordinal_stage, y=Tissue_ForDevelopementalAnalysis, color=DifferentialSplicingAndExpressionGroup)) +
  geom_point(aes(size=n)) +
  facet_wrap(~ID_Species, scales="free_x")

ManuallyAnnotated_Neonate_v_Adults %>%
  filter(!is.na(DifferentialSplicingAndExpressionGroup)) %>%
  count(ID_Species, Tissue_ForDevelopementalAnalysis, DifferentialSplicingAndExpressionGroup) %>%
  ggplot(aes(x=Tissue_ForDevelopementalAnalysis, fill=DifferentialSplicingAndExpressionGroup, y=n)) +
  geom_col(position=position_dodge()) +
  facet_wrap(~ID_Species) +
  labs(y="Sample size") +
  Rotate_x_labels
```
Ok, so for alsmost all species:tissue pairs, I have a reasonably sized sample (n>=4) for the neonate and the adult group. Now I want to make contrasts for each adult tissue pair within species, and neonate vs adult within each tissue:species pair.

Let's rewrite the Contrasts variable to include all of these

```{r}
Approach6.samples.grouped <- ManuallyAnnotated_Neonate_v_Adults %>%
  filter(!is.na(DifferentialSplicingAndExpressionGroup)) %>%
  mutate(Group = recode(DifferentialSplicingAndExpressionGroup, "neonate"="Early", "adult"="Late")) %>%
  mutate(Group = factor(Group)) %>%
  dplyr::select(Group, Tissue_ForDevelopementalAnalysis, ID, ID_Species)

bind_rows(
  list(
    "BeforeM_vs_AfterS" = Approach1.samples.grouped,
    "BeforeB_vs_AfterB" = Approach2.samples.grouped,
    "BeforeB_vs_AferS" = Approach3.samples.grouped,
    "FirstHalf_vs_SecondHalf" = Approach4.samples.grouped,
    "First10_vs_Second10" = Approach5.samples.grouped,
    "Neonate_vs_Adult" = Approach6.samples.grouped
  ),
  .id="Approach"
) %>%
  count(Group, Tissue_ForDevelopementalAnalysis, Approach, ID_Species) %>%
  ggplot(aes(x=Tissue_ForDevelopementalAnalysis, y=n, fill=Group)) +
  geom_col(position='stack') +
  facet_grid(Approach~ID_Species) +
  Rotate_x_labels

Contrasts <- bind_rows(
  list(
    "BeforeM_vs_AfterS" = Approach1.samples.grouped,
    "BeforeB_vs_AfterB" = Approach2.samples.grouped,
    "BeforeB_vs_AferS" = Approach3.samples.grouped,
    "FirstHalf_vs_SecondHalf" = Approach4.samples.grouped,
    "First10_vs_Second10" = Approach5.samples.grouped,
    "Neonate_vs_Adult" = Approach6.samples.grouped
  ),
  .id="Approach"
) %>%
  add_count(Group, Tissue_ForDevelopementalAnalysis, Approach, ID_Species) %>%
  group_by(Group, Tissue_ForDevelopementalAnalysis, Approach, ID_Species) %>%
  filter(all(n>=2)) %>%
  ungroup() %>%
  group_by(Tissue_ForDevelopementalAnalysis, Approach, ID_Species) %>%
  filter(any(Group=="Late") & any(Group=="Early")) %>%
  ungroup() %>%
  mutate(GenomeName = recode(ID_Species, "Chicken"="Chicken_UCSC.galGal6_ensv101", "Rabbit"="Rabbit_UCSC.oryCun2_ensv101", "Opossum"="Opossum_UCSC.monDom5_ensvUnknown", "Human"="Human_UCSC.hg38_GencodeComprehensive46", "Mouse"="Mouse_UCSC.mm39_GencodeComprehensive46", "Rat"="Rat_UCSC.rn7_RefSeqv108", "Macaque"="Macaque_UCSC.rheMac10_ensv101")) %>%
  mutate(ContrastName = str_glue("{Approach}.{Tissue_ForDevelopementalAnalysis}.{ID_Species}"))


Contrasts %>%
  distinct(ContrastName, GenomeName) %>%
  dplyr::select(ContrastName, GenomeName) %>%
  write_tsv("../code/config/CordosoTimeSeriesContrasts.tsv")

dir.create("../code/MazinLeafcutterAnalysis/ContrastGroupFiles/")

Contrasts %>%
  dplyr::select(ContrastName, ID, ID_Species, Tissue_ForDevelopementalAnalysis, Group) %>%
  arrange(ID_Species, Tissue_ForDevelopementalAnalysis, Group) %>%
  group_by(ContrastName) %>%
  dplyr::select(ID, Group) %>%
  group_walk(~ write_delim(.x, paste0("../code/MazinLeafcutterAnalysis/ContrastGroupFiles/", .y$ContrastName, ".txt"), delim = ' ', col_names = F))
```

Now let's add the constrasts for adult tissue vs adult tissue. Let's write this to a different set of contrasts.

```{r}

AdultTissueContrasts <- full_join(
  ManuallyAnnotated_Neonate_v_Adults %>%
    filter(DifferentialSplicingAndExpressionGroup == "adult") %>%
    dplyr::select(ID, ID_Species, Tissue_ForDevelopementalAnalysis) %>%
    distinct(ID_Species, Tissue_ForDevelopementalAnalysis),
  ManuallyAnnotated_Neonate_v_Adults %>%
    filter(DifferentialSplicingAndExpressionGroup == "adult") %>%
    dplyr::select(ID, ID_Species, Tissue_ForDevelopementalAnalysis) %>%
    distinct(ID_Species, Tissue_ForDevelopementalAnalysis),
  by=c("ID_Species")
) %>%
  filter(!Tissue_ForDevelopementalAnalysis.x==Tissue_ForDevelopementalAnalysis.y) %>%
  mutate(OrderedComparison = paste(pmin(Tissue_ForDevelopementalAnalysis.x, Tissue_ForDevelopementalAnalysis.y), 
                        pmax(Tissue_ForDevelopementalAnalysis.x, Tissue_ForDevelopementalAnalysis.y), sep= "-")) %>%
  distinct(ID_Species, OrderedComparison, .keep_all=T) %>%
  pivot_longer(c("Tissue_ForDevelopementalAnalysis.x", "Tissue_ForDevelopementalAnalysis.y"), names_to = "X_or_Y", values_to = "Tissue_ForDevelopementalAnalysis") %>%
  dplyr::select(-X_or_Y) %>%
  inner_join(
    ManuallyAnnotated_Neonate_v_Adults %>%
      filter(DifferentialSplicingAndExpressionGroup == "adult") %>%
      dplyr::select(ID, ID_Species, Tissue_ForDevelopementalAnalysis) 
  ) %>%
  mutate(Tissue_ForDevelopementalAnalysis = factor(Tissue_ForDevelopementalAnalysis, levels=c("Brain", "Cerebellum", "Heart", "Kidney","Liver", "Ovary", "Testis"))) %>%
  arrange(ID_Species, OrderedComparison, Tissue_ForDevelopementalAnalysis, ID) %>%
  mutate(ContrastName = str_glue("{ID_Species}_{OrderedComparison}")) %>%
  inner_join(
    read_tsv("../code/config/samples.tsv") %>%
      dplyr::select(ID=SampleName, GenomeName = STARGenomeName)
  )


AdultTissueContrasts %>%
  distinct(ContrastName, GenomeName) %>%
  dplyr::select(ContrastName, GenomeName) %>%
  write_tsv("../code/config/CordosoTissueContrasts.tsv")

dir.create("../code/MazinLeafcutterAnalysis/ContrastAdultTissueGroupFiles/")

AdultTissueContrasts %>%
  dplyr::select(ContrastName, ID, Tissue_ForDevelopementalAnalysis) %>%
  arrange(ContrastName, Tissue_ForDevelopementalAnalysis) %>%
  group_by(ContrastName) %>%
  dplyr::select(ID, Group=Tissue_ForDevelopementalAnalysis) %>%
  group_walk(~ write_delim(.x, paste0("../code/MazinLeafcutterAnalysis/ContrastAdultTissueGroupFiles/", .y$ContrastName, ".txt"), delim = ' ', col_names = F))

```

