---
title: "2024-09-09_Organize_ConserveddevASJuncs"
output: html_document
date: '2024-09-09'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

## Intro

```{r}
library(tidyverse)
library(data.table)

# Set theme
theme_set(
  theme_classic() +
  theme(text=element_text(size=16,  family="Helvetica")))

# I use layer a lot, to rotate long x-axis labels
Rotate_x_labels <- theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```

First, as a sanity check, let's look at log2FC by classification status in human based on naRNA-seq data...

```{r}

dat.KD <- fread("/project2/yangili1/cfbuenabadn/ChromatinSplicingQTLs/code/SplicingAnalysis/CombinedJuncTables/NMD_KD.tsv.gz")


juncs.long.summary <- 
  dat.KD %>%
  dplyr::select(chrom, start, stop, strand, Dataset, Count) %>%
  group_by(Dataset, chrom, start, stop) %>%
  summarise(Sum=sum(Count)) %>%
  ungroup()

Classifications <- read_tsv("../code/MazinLeafcutterAnalysis/ClassifyJuncs/Human_UCSC.hg38_GencodeComprehensive46.AllObserved._junction_classifications.txt") %>%
  separate(Intron_coord, into=c("chrom", "start", "stop"), sep="[:-]", convert=T, remove=F)

Classifications %>%
  head()

juncs.long.summary %>%
  distinct(Dataset)

Joined <- juncs.long.summary %>%
  filter(Dataset %in% c("HeLa.dKD", "HeLa.scr")) %>%
  group_by(Dataset) %>%
  mutate(RPM = Sum/sum(Sum)*1E6) %>%
  ungroup() %>%
  pivot_wider(names_from = c("Dataset"), values_from=c("Sum", "RPM")) %>%
  inner_join(
    Classifications
    )

min(Joined$RPM_HeLa.dKD, na.rm=T)
min(Joined$RPM_HeLa.scr, na.rm=T)


Joined %>%
  replace_na(list(Sum_HeLa.scr=0, Sum_HeLa.dKD=0, RPM_HeLa.dKD=1E-4, RPM_HeLa.scr=1E-4)) %>%
  mutate(DeltaRPM = log2(RPM_HeLa.dKD/RPM_HeLa.scr)) %>%
  mutate(Annot = if_else(Annot, "Annot", "Unannot")) %>%
  mutate(Coding = if_else(Coding, "Productive", "Unproductive")) %>%
  mutate(Group = paste(Coding, Annot)) %>%
  ggplot(aes(x=DeltaRPM, color=Group)) +
    stat_ecdf() +
    coord_cartesian(xlim=c(-5,5)) +
  geom_vline(xintercept = 0)

```


```{r}
Classifications.clustered <- read_tsv("../code/MazinLeafcutterAnalysis/ClassifyJuncs/Human_UCSC.hg38_GencodeComprehensive46.Clustered._junction_classifications.txt") %>%
  separate(Intron_coord, into=c("chrom", "start", "stop"), sep="[:-]", convert=T, remove=F)

Classifications.clustered %>%
  arrange(chrom, start, stop)

Classifications.clustered %>%
  count(Annot, Coding)

Joined.clustered <- juncs.long.summary %>%
  filter(Dataset %in% c("HeLa.dKD", "HeLa.scr")) %>%
  group_by(Dataset) %>%
  mutate(RPM = Sum/sum(Sum)*1E6) %>%
  ungroup() %>%
  pivot_wider(names_from = c("Dataset"), values_from=c("Sum", "RPM")) %>%
  inner_join(
    Classifications.clustered
    )

Joined.clustered %>%
  count(Annot, Coding)

Classifications %>%
  inner_join(Classifications.clustered, by=c("Intron_coord", "Gene_name")) %>%
  count(Annot.x, Annot.y, Coding.x, Coding.y)

Classifications %>%
  inner_join(Classifications.clustered, by=c("Intron_coord", "Gene_name")) %>%
  count(Annot.x, Annot.y, Coding.x, Coding.y) %>%
  mutate(Coding.x = if_else(Coding.x, "AllObs_Productive", "AllObs_Unproductive")) %>%
  mutate(Coding.y = if_else(Coding.y, "Clustered_Productive", "Clustered_Unproductive")) %>%
  mutate(Annot = if_else(Annot.x, "Annotated", "Unannotated")) %>%
  mutate(CodingGroup = str_glue("{Coding.x} {Coding.y}")) %>%
  ggplot(aes(x=Annot, y=n, fill=CodingGroup)) +
  geom_col(position='stack')


Joined.clustered %>%
  replace_na(list(Sum_HeLa.scr=0, Sum_HeLa.dKD=0, RPM_HeLa.dKD=1E-4, RPM_HeLa.scr=1E-4)) %>%
  mutate(DeltaRPM = log2(RPM_HeLa.dKD/RPM_HeLa.scr)) %>%
  mutate(Annot = if_else(Annot, "Annot", "Unannot")) %>%
  mutate(Coding = if_else(Coding, "Productive", "Unproductive")) %>%
  mutate(Group = paste(Coding, Annot)) %>%
  ggplot(aes(x=DeltaRPM, color=Group)) +
    stat_ecdf() +
    coord_cartesian(xlim=c(-5,5)) +
  geom_vline(xintercept = 0)

```



I now understand the problem... I did not format the input correctly... Non-clustered juncs were right, but not clusterd juncs. So I get all kinds of weird results...


I have fixed that problem.. But now I still suspect an issue in the script. Let's clone Chao's most recent version of the script, and run the forward classifier script.

```{r}
Classifications.100K.All <- read_tsv("../code/scratch/Leaf2Clustered_junction_classifications.txt") %>%
  separate(Intron_coord, into=c("chrom", "start", "stop"), sep="[:-]", convert=T, remove=F)

Classifications.100K.All %>%
  count(Annot, Coding)

Joined.clustered <- juncs.long.summary %>%
  filter(Dataset %in% c("HeLa.dKD", "HeLa.scr")) %>%
  group_by(Dataset) %>%
  mutate(RPM = Sum/sum(Sum)*1E6) %>%
  ungroup() %>%
  pivot_wider(names_from = c("Dataset"), values_from=c("Sum", "RPM")) %>%
  inner_join(
    Classifications.100K.All %>%
      mutate(stop = as.numeric(stop)) %>%
      mutate(stop = if_else(Annot, stop, stop))
    )

Joined.clustered %>%
  count(Annot, Coding)

Joined.clustered %>%
  replace_na(list(Sum_HeLa.scr=0, Sum_HeLa.dKD=0, RPM_HeLa.dKD=1E-4, RPM_HeLa.scr=1E-4)) %>%
  mutate(DeltaRPM = log2(RPM_HeLa.dKD/RPM_HeLa.scr)) %>%
  mutate(Annot = if_else(Annot, "Annot", "Unannot")) %>%
  mutate(Coding = if_else(Coding, "Productive", "Unproductive")) %>%
  mutate(Group = paste(Coding, Annot)) %>%
  ggplot(aes(x=DeltaRPM, color=Group)) +
    stat_ecdf() +
    coord_cartesian(xlim=c(-5,5)) +
  geom_vline(xintercept = 0)

```
Wow. that is clearly much better. Ok. so I dunno what the bug was. but now it's fixed.
